---
title: "Vaisala vs Arduino Data Comparison"
author: "Leona Neftaliem"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE)
```

# Libraries
```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
library(ggplot2)
library(scales)
library(Metrics)
library(ranger)      # Random forest models

# Helper function to parse incoming ET/EST timestamps and convert them to UTC
parse_to_utc <- function(ts, format = NULL) {
  if (is.null(ts)) return(NULL)
  
  # If it is already a POSIXct/POSIXlt object
  if (inherits(ts, "POSIXt")) {
    # If the timezone is already UTC, return it. Otherwise, force to America/New_York and convert to UTC
    if (identical(tz(ts), "UTC")) {
      return(ts)
    } else {
      return(lubridate::with_tz(lubridate::force_tz(ts, "America/New_York"), "UTC"))
    }
  }
  
  # Handle character/factor timestamps
  ts_str <- as.character(ts)
  
  # Clean up ISO 8601 formatting like 'Z' or 'T' if present
  # because all incoming data is in ET/EST, even if mislabeled with a Z
  ts_str <- gsub("Z$", "", ts_str)
  ts_str <- gsub("T", " ", ts_str)
  
  # Try parsing with format if provided
  if (!is.null(format)) {
    parsed <- as.POSIXct(ts_str, format = format, tz = "America/New_York")
    return(lubridate::with_tz(parsed, "UTC"))
  }
  
  # Try parsing with common formats
  formats <- c("%Y-%m-%d %H:%M:%S", "%m/%d/%Y %H:%M:%S", "%Y-%m-%d %H:%M", "%m/%d/%Y %H:%M", "%Y-%m-%d", "%m/%d/%Y")
  for (fmt in formats) {
    parsed <- as.POSIXct(ts_str, format = fmt, tz = "America/New_York")
    if (!all(is.na(parsed))) {
      return(lubridate::with_tz(parsed, "UTC"))
    }
  }
  
  # Fallback to general parsing
  parsed <- as.POSIXct(ts_str, tz = "America/New_York")
  return(lubridate::with_tz(parsed, "UTC"))
}
```

# Overview

This script builds calibration models to correct low-cost Arduino sensor readings 
using co-located Vaisala reference data. It produces regression equations that can 
be applied to any Arduino sensor deployment.

## How Calibration Works

1. **Co-locate** an Arduino sensor next to a Vaisala reference monitor
2. **Match** their readings by timestamp (rounded to the nearest hour)
3. **Fit** a regression model: `Vaisala = intercept + slope×Arduino + slope×humidity + slope×temp`
4. **Apply** the model to correct future Arduino readings

## Data Sources

- **General calibration**: Built from the original AQM_206 co-location data 
  (`AQM_206_AQM_ArduinoTable.dat`, `AQM_206_AQM_ArduinoTable2.dat`, `AQM_206_AQM_VaisalaTable.dat`)
- **Site-specific calibrations**: Built from SERC calibration CSVs 
  (`Strength2LoveFarm_SERCCal.csv`, `BayBrookElementary_SERCCal.csv`, `VaisalaTable.csv`)

## Minimum Data Requirements

All calibrations use **hourly** matched data, so **1 day = 24 data points**.

- **To fit (create) a model**: Need at least **24 hourly observations** (1 day of co-location)
- **To apply an existing model**: No minimum—works on any amount of data

This means if you have a short deployment (e.g., 6-7 days at the Baltimore dock), 
you can either:
1. **Apply** the pre-existing general calibration to your data (no minimum needed), or
2. **Fit** a new site-specific model if you have at least 1 day of co-located reference data

# Calibration Settings
```{r calibration-settings}
# Minimum number of HOURLY data points needed to fit a calibration model.
# 24 hours = 1 day of co-location. This applies to both general and site-specific models.
# Applying an existing model has no minimum—it works on any amount of data.
min_hours_for_calibration <- 24

# Helper function to handle predictions for both RF and LM 
predict_best_model <- function(mi, newdata) {
  if (is.null(mi) || is.null(mi$model)) return(NULL)
  
  m_type <- if(!is.null(mi$model_type)) mi$model_type else "rf"
  
  if (m_type == "lm") {
    # Linear Model prediction
    return(as.numeric(predict(mi$model, newdata = newdata)))
  } else if (m_type == "bias_only") {
    # Bias-correction (Mean shift)
    return(rep(as.numeric(mi$bias_value), nrow(newdata)))
  } else {
    # Random Forest (ranger) prediction
    return(as.numeric(predict(mi$model, data = newdata)$predictions))
  }
}

# Sensors that have lab co-location data and are eligible for the
# site-specific second-stage linear correction (applied from 2026-12-01 onward).
# Fill in the actual sensor IDs that appear in your deployment data.
# Example: calibrated_sensors <- c("SENSOR_A", "SENSOR_B")
calibrated_sensors <- character(0)   # <-- populate with your sensor IDs

# Date after which site-specific correction is applied to calibrated sensors.
# Before this date all sensors use the general RF correction.
calibration_cutoff_date <- as.Date("2025-12-01")

cat("Calibration settings:\n")
cat("  Minimum hours for fitting a model:", min_hours_for_calibration,
    "(", min_hours_for_calibration/24, "day(s) of co-location)\n")
cat("  Calibration cutoff date:", as.character(calibration_cutoff_date), "\n")
cat("  Sensors with lab data (calibrated_sensors):",
    if(length(calibrated_sensors) == 0) "<none set>" else paste(calibrated_sensors, collapse = ", "), "\n")
```

# Data Loading
```{r}
# Place raw co-location files (see data/README.md) in data/current_data/,
# and site-specific calibration folders in data/Cal_Stations/
data_dir <- "data/current_data"

# Load the three data files
arduino1_file <- file.path(data_dir, "AQM_206_AQM_ArduinoTable.dat")
arduino2_file <- file.path(data_dir, "AQM_206_AQM_ArduinoTable2.dat")
vaisala_file <- file.path(data_dir, "AQM_206_AQM_VaisalaTable.dat")

# Check if files exist
has_data <- any(file.exists(c(arduino1_file, arduino2_file, vaisala_file)))
cat("Has data files:", has_data, "\n")
```

# -------------------------------------------------------------------------
# DATA PROCESSING & INGESTION
# -------------------------------------------------------------------------
# Load raw sensor and reference (Vaisala) data. 
# Handles multiple encodings and provides sample data fallbacks.
# -------------------------------------------------------------------------
```{r}
# Define column structures
vaisala_columns <- c("timestamp", "record", "vaisalaString", "no2_ppb", "co_ppb", "no_ppb", "o3_ppb", "pm1_ugm3", "pm25_ugm3", "pm10_ugm3", "air_pres_pa", "air_temp_c", "humidity_perc_rh")
arduino_columns <- c("timestamp", "record", "arduinostring", "rtctime", "sl_id", "program", "ard_temp_c", "ard_press_pa", "ard_humidity_perc_rh", "ard_voc", "ard_alt", "ard_co2_ppm", "ard_o3_ppb", "ard_pm1_1_ugm3", "ard_pm25_1_ugm3", "ard_pm10_1_ugm3", "ard_pm1_2_ugm3", "ard_pm25_2_ugm3", "ard_pm10_2_ugm3")

# Load data or generate sample data
if(has_data) {
  cat("Loading real data files...\n")
  
  # Function to safely read CSV files with encoding handling
  safe_read_csv <- function(file_path, columns, skip_lines = 3) {
    if(!file.exists(file_path)) {
      cat("File does not exist:", basename(file_path), "\n")
      return(NULL)
    }
    
    # Try different encodings
    encodings <- c("UTF-8", "latin1", "ISO-8859-1", "windows-1252")
    
    for(enc in encodings) {
      tryCatch({
        # Read as character first to avoid type conversion errors
        data <- read.csv(file_path, header=TRUE, skip=skip_lines, 
                        fileEncoding=enc, stringsAsFactors=FALSE,
                        check.names=FALSE, quote="\"", comment.char="",
                        as.is=TRUE)
        
        # Verify we got data
        if(nrow(data) > 0 && ncol(data) > 0) {
          colnames(data) <- columns
          cat("Successfully loaded", basename(file_path), "with encoding:", enc, "\n")
          return(data)
        }
      }, error = function(e) {
        # Try next encoding
        return(NULL)
      })
    }
    
    # If all encodings fail, try without specifying encoding
    tryCatch({
      data <- read.csv(file_path, header=TRUE, skip=skip_lines, 
                      stringsAsFactors=FALSE, check.names=FALSE,
                      fileEncoding="", quote="\"", comment.char="",
                      as.is=TRUE)
      colnames(data) <- columns
      cat("Successfully loaded", basename(file_path), "with default encoding\n")
      return(data)
    }, error = function(e) {
      cat("Error loading", basename(file_path), ":", e$message, "\n")
      return(NULL)
    })
  }
  
  arduino1_data <- safe_read_csv(arduino1_file, arduino_columns)
  arduino2_data <- safe_read_csv(arduino2_file, arduino_columns)
  vaisala_data <- safe_read_csv(vaisala_file, vaisala_columns)
  
  # Check if any files failed to load
  if(is.null(arduino1_data) || is.null(arduino2_data) || is.null(vaisala_data)) {
    cat("Warning: Some files failed to load. Falling back to sample data generation.\n")
    has_data <- FALSE
  }
}

# Generate sample data if files don't exist or failed to load
if(!has_data) {
  cat("Generating sample data for demonstration...\n")
  
  # Generate sample data
  set.seed(123)
  n_points <- 1000
  
  # Create sample timestamps in UTC (converted from America/New_York)
  start_time <- parse_to_utc("2024-01-01 00:00:00")
  timestamps <- start_time + seq(0, n_points * 3600, by = 3600)  # Hourly data
  
  # Generate sample Arduino data
  arduino1_data <- data.frame(
    timestamp = timestamps[1:n_points],
    record = 1:n_points,
    arduinostring = paste0("arduino1_", 1:n_points),
    rtctime = timestamps[1:n_points],
    sl_id = rep("A001", n_points),
    program = rep("AQM", n_points),
    ard_temp_c = rnorm(n_points, 20, 5),
    ard_press_pa = rnorm(n_points, 101325, 1000),
    ard_humidity_perc_rh = rnorm(n_points, 50, 15),
    ard_voc = rnorm(n_points, 100, 20),
    ard_alt = rnorm(n_points, 100, 10),
    ard_co2_ppm = rnorm(n_points, 400, 50),
    ard_o3_ppb = rnorm(n_points, 30, 10),
    ard_pm1_1_ugm3 = rnorm(n_points, 5, 2),
    ard_pm25_1_ugm3 = rnorm(n_points, 10, 5),
    ard_pm10_1_ugm3 = rnorm(n_points, 15, 8),
    ard_pm1_2_ugm3 = rnorm(n_points, 5, 2),
    ard_pm25_2_ugm3 = rnorm(n_points, 10, 5),
    ard_pm10_2_ugm3 = rnorm(n_points, 15, 8)
  )
  
  arduino2_data <- arduino1_data
  arduino2_data$arduinostring <- paste0("arduino2_", 1:n_points)
  
  # Generate sample Vaisala data
  vaisala_data <- data.frame(
    timestamp = timestamps[1:n_points],
    record = 1:n_points,
    vaisalaString = paste0("vaisala_", 1:n_points),
    no2_ppb = rnorm(n_points, 20, 5),
    co_ppb = rnorm(n_points, 1, 0.5),
    no_ppb = rnorm(n_points, 5, 2),
    o3_ppb = rnorm(n_points, 35, 8),
    pm1_ugm3 = rnorm(n_points, 6, 2),
    pm25_ugm3 = rnorm(n_points, 12, 4),
    pm10_ugm3 = rnorm(n_points, 18, 6),
    air_pres_pa = rnorm(n_points, 101325, 1000),
    air_temp_c = rnorm(n_points, 20, 5),
    humidity_perc_rh = rnorm(n_points, 50, 15)
  )
}

# Convert timestamps to UTC (from America/New_York) and add unit identifiers
arduino1_data$timestamp <- parse_to_utc(arduino1_data$timestamp, format = "%Y-%m-%d %H:%M:%S")
arduino2_data$timestamp <- parse_to_utc(arduino2_data$timestamp, format = "%Y-%m-%d %H:%M:%S")
vaisala_data$timestamp <- parse_to_utc(vaisala_data$timestamp, format = "%Y-%m-%d %H:%M:%S")

arduino1_data$datetime <- arduino1_data$timestamp
arduino2_data$datetime <- arduino2_data$timestamp
vaisala_data$datetime <- vaisala_data$timestamp

arduino1_data$unit <- "Arduino_1"
arduino2_data$unit <- "Arduino_2" 
vaisala_data$unit <- "Vaisala"

# Clean data: Convert NaN and -7999 to NA, then remove rows with NA
arduino1_clean <- arduino1_data %>%
  mutate(across(everything(), ~ {
    if (is.character(.)) {
      x <- na_if(., "NAN")
      x[x == "-7999"] <- NA
      x
    } else if (is.numeric(.)) {
      x <- .
      x[is.nan(x) | x == -7999] <- NA
      x
    } else {
      .
    }
  })) %>%
  drop_na()

arduino2_clean <- arduino2_data %>%
  mutate(across(everything(), ~ {
    if (is.character(.)) {
      x <- na_if(., "NAN")
      x[x == "-7999"] <- NA
      x
    } else if (is.numeric(.)) {
      x <- .
      x[is.nan(x) | x == -7999] <- NA
      x
    } else {
      .
    }
  })) %>%
  drop_na()

vaisala_clean <- vaisala_data %>%
  mutate(across(everything(), ~ {
    # If the column is character, replace "NaN" string with NA
    if (is.character(.)) {
      na_if(., "NAN")
    # If numeric, replace actual NaN values with NA
    } else if (is.numeric(.)) {
      x <- .
      x[is.nan(x)] <- NA
      x
    # Otherwise, leave as is
    } else {
      .
    }
  })) %>%
  drop_na() 

cat("Data cleaning completed:\n")
cat("- Arduino1: ", nrow(arduino1_data), " -> ", nrow(arduino1_clean), " rows\n")
cat("- Arduino2: ", nrow(arduino2_data), " -> ", nrow(arduino2_clean), " rows\n")
cat("- Vaisala: ", nrow(vaisala_data), " -> ", nrow(vaisala_clean), " rows\n")
```


# -------------------------------------------------------------------------
# DATA PREPARATION & ALIGNMENT
# -------------------------------------------------------------------------
# Aligns Arduino and Vaisala timestamps to create a synchronized 
# training dataset (all_data_hourly).
# -------------------------------------------------------------------------
```{r}
# Create averaged Arduino data before developing corrections
cat("Creating averaged Arduino dataset for more robust corrections...\n")


# Ensure both datasets have POSIXct timestamp columns in UTC
arduino1_data$timestamp <- parse_to_utc(arduino1_data$timestamp)
arduino2_data$timestamp <- parse_to_utc(arduino2_data$timestamp)
  
# Prepare Arduino data columns for averaging
arduino1_prep <- arduino1_clean %>%
  dplyr::select(timestamp, ard_o3_ppb, ard_pm1_1_ugm3, ard_pm25_1_ugm3, ard_pm10_1_ugm3, ard_temp_c, ard_humidity_perc_rh)

arduino2_prep <- arduino2_clean %>%
  dplyr::select(timestamp, ard_o3_ppb, ard_pm1_1_ugm3, ard_pm25_1_ugm3, ard_pm10_1_ugm3, ard_temp_c, ard_humidity_perc_rh)

# Merge Arduino datasets by timestamp (full join so it works even if one unit is empty)
arduino_merged <- full_join(
  arduino1_prep,
  arduino2_prep,
  by = "timestamp",
  suffix = c("_a1", "_a2")
)

# Average the two Arduino units for each pollutant (falls back to the one available)
arduino_avg <- arduino_merged %>%
  mutate(
    o3_ppm = rowMeans(cbind(as.numeric(ard_o3_ppb_a1), as.numeric(ard_o3_ppb_a2)), na.rm = TRUE) / 1000,
    pm1_ugm3 = rowMeans(cbind(as.numeric(ard_pm1_1_ugm3_a1), as.numeric(ard_pm1_1_ugm3_a2)), na.rm = TRUE),
    pm25_ugm3 = rowMeans(cbind(as.numeric(ard_pm25_1_ugm3_a1), as.numeric(ard_pm25_1_ugm3_a2)), na.rm = TRUE),
    pm10_ugm3 = rowMeans(cbind(as.numeric(ard_pm10_1_ugm3_a1), as.numeric(ard_pm10_1_ugm3_a2)), na.rm = TRUE),
    temp_c = rowMeans(cbind(as.numeric(ard_temp_c_a1), as.numeric(ard_temp_c_a2)), na.rm = TRUE),
    humidity_perc_rh = rowMeans(cbind(as.numeric(ard_humidity_perc_rh_a1), as.numeric(ard_humidity_perc_rh_a2)), na.rm = TRUE),
    unit = "Arduino_Averaged"
  ) %>%
  dplyr::select(timestamp, o3_ppm, pm1_ugm3, pm25_ugm3, pm10_ugm3, temp_c, humidity_perc_rh, unit)


# Prepare Vaisala data
vaisala_combined <- vaisala_clean %>%
  dplyr::select(timestamp, record, o3_ppb, pm1_ugm3, pm25_ugm3, pm10_ugm3, air_temp_c, humidity_perc_rh) %>%
  mutate(
    unit = "Vaisala", 
    timestamp_rounded = timestamp,
    temp_c = as.numeric(air_temp_c),  # Convert to numeric
    o3_ppm = as.numeric(o3_ppb) / 1000,
    pm1_ugm3 = as.numeric(pm1_ugm3),
    pm25_ugm3 = as.numeric(pm25_ugm3),
    pm10_ugm3 = as.numeric(pm10_ugm3),
    humidity_perc_rh = as.numeric(humidity_perc_rh)
  ) %>%
  dplyr::select(-air_temp_c, -o3_ppb)  # Remove the old column name


# Combine data using inner join to match timestamps
all_data <- inner_join(arduino_avg, vaisala_combined, by = "timestamp", suffix = c("_arduino", "_vaisala")) %>%
  dplyr::select(
    timestamp,
    # Arduino measurements
    o3_ppm_arduino, pm1_ugm3_arduino, pm25_ugm3_arduino, pm10_ugm3_arduino,
    temp_c_arduino, humidity_perc_rh_arduino,
    # Vaisala measurements  
    o3_ppm_vaisala, pm1_ugm3_vaisala, pm25_ugm3_vaisala, pm10_ugm3_vaisala,
    temp_c_vaisala, humidity_perc_rh_vaisala
  )

# Aggregate to HOURLY data (for consistent calibration approach)
cat("Aggregating data to hourly level...\n")

# Ensure timestamp is POSIXct in UTC
all_data <- all_data %>%
  mutate(timestamp = parse_to_utc(timestamp))

all_data_hourly <- all_data %>%
  mutate(timestamp_hour = floor_date(timestamp, unit = "hour")) %>%
  group_by(timestamp_hour) %>%
  summarise(
    # Arduino measurements - hourly means
    o3_ppm_arduino = mean(o3_ppm_arduino, na.rm = TRUE),
    pm1_ugm3_arduino = mean(pm1_ugm3_arduino, na.rm = TRUE),
    pm25_ugm3_arduino = mean(pm25_ugm3_arduino, na.rm = TRUE),
    pm10_ugm3_arduino = mean(pm10_ugm3_arduino, na.rm = TRUE),
    temp_c_arduino = mean(temp_c_arduino, na.rm = TRUE),
    humidity_perc_rh_arduino = mean(humidity_perc_rh_arduino, na.rm = TRUE),
    # Vaisala measurements - hourly means
    o3_ppm_vaisala = mean(o3_ppm_vaisala, na.rm = TRUE),
    pm1_ugm3_vaisala = mean(pm1_ugm3_vaisala, na.rm = TRUE),
    pm25_ugm3_vaisala = mean(pm25_ugm3_vaisala, na.rm = TRUE),
    pm10_ugm3_vaisala = mean(pm10_ugm3_vaisala, na.rm = TRUE),
    temp_c_vaisala = mean(temp_c_vaisala, na.rm = TRUE),
    humidity_perc_rh_vaisala = mean(humidity_perc_rh_vaisala, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  rename(timestamp = timestamp_hour)

# Report dataset summary
n_hours <- nrow(all_data_hourly)
n_days <- n_hours / 24
date_range <- range(as.Date(all_data_hourly$timestamp))

cat("\n--- General Calibration Data Summary ---\n")
cat("Total hourly observations:", n_hours, "\n")
cat("Equivalent days:", round(n_days, 1), "\n")
cat("Date range:", as.character(date_range[1]), "to", as.character(date_range[2]), "\n")

if(n_hours >= min_hours_for_calibration) {
  cat("Status: SUFFICIENT for calibration (", n_hours, " >= ", min_hours_for_calibration, " hours)\n", sep = "")
} else {
  cat("Status: INSUFFICIENT for calibration (", n_hours, " < ", min_hours_for_calibration, " hours)\n", sep = "")
}

# Use hourly data for modeling
all_data <- all_data_hourly

# --- Step 2: Hourly observation warning for calibrated sensors ---
# Flag any calibrated sensor whose colocation period has < 72 hourly observations.
# 72 hours = 3 full days — a practical minimum for reliable RF training.
if(length(calibrated_sensors) > 0) {
  cat("\n--- Calibrated Sensor Hourly Observation Check ---\n")
  # The general all_data represents the colocation period for the general model.
  # Site-specific sensors are checked again inside create_site_calibration_models.
  cat("Calibrated sensors will be checked for >= 72 hourly obs in their site-specific colocation data.\n")
  cat("(General colocation hourly obs:", n_hours, ")\n")
}

```

# General Calibration Model
```{r general-model}
# General Model Creation (Stage 1)
# Ozone: Forced MLR for stability on flat signals.
# PM: Random Forest for non-linear RH/Temp effects.
create_general_calibration_models <- function(all_data) {
  if(nrow(all_data) == 0) {
    cat("No data available for modeling.\n")
    return(NULL)
  }

  target_params  <- c("o3_ppm", "pm1_ugm3", "pm25_ugm3", "pm10_ugm3")
  general_models <- list()

  for(param in target_params) {
    arduino_col <- paste0(param, "_arduino")
    vaisala_col <- paste0(param, "_vaisala")

    if(!(arduino_col %in% colnames(all_data)) || !(vaisala_col %in% colnames(all_data))) {
      cat("Missing columns for", param, "\n")
      next
    }

    model_data <- all_data %>%
      filter(
        !is.na(.data[[arduino_col]]), !is.na(.data[[vaisala_col]]),
        is.finite(.data[[arduino_col]]), is.finite(.data[[vaisala_col]]),
        !is.na(temp_c_arduino), !is.na(humidity_perc_rh_arduino),
        is.finite(temp_c_arduino), is.finite(humidity_perc_rh_arduino)
      )

    if(nrow(model_data) < min_hours_for_calibration) {
      cat("Insufficient data for", param, "(", nrow(model_data), "hours; need ",
          min_hours_for_calibration, ")\n")
      next
    }

    # 80/20 random stratified split (replacing temporal split)
    set.seed(42)
    n_train <- floor(0.8 * nrow(model_data))
    train_indices <- sample(seq_len(nrow(model_data)), n_train)
    train_df <- model_data[train_indices, ]
    test_df  <- model_data[-train_indices, ]

    # --- Diagnostics (per user request) ---
    cat(sprintf("\n--- Diagnostics for %s ---\n", param))
    cat("  Training date range:", as.character(min(as.Date(train_df$timestamp))), 
        "to", as.character(max(as.Date(train_df$timestamp))), "\n")
    cat("  Unique dates in training:", length(unique(as.Date(train_df$timestamp))), "\n")
    
    # Correlation check (verify alignment)
    cor_val <- cor(train_df[[arduino_col]], train_df[[vaisala_col]], use = "complete.obs")
    cat(sprintf("  Correlation (%s vs %s): %.3f\n", arduino_col, vaisala_col, cor_val))

    rf_formula <- as.formula(paste0(
      vaisala_col, " ~ ", arduino_col, " + humidity_perc_rh_arduino + temp_c_arduino"
    ))

    # Fit models (optimized for speed: skip RF for Ozone)
    ss_tot   <- sum((test_df[[vaisala_col]] - mean(test_df[[vaisala_col]]))^2)
    rf_preds <- NULL
    
    if (ss_tot > 0) {
      if (param != "o3_ppb") {
        rf_model <- ranger(
          formula   = rf_formula,
          data      = train_df,
          importance = "impurity",
          num.trees  = 500,
          seed       = 42,
          num.threads = parallel::detectCores() # Speed optimization
        )
        rf_preds <- predict(rf_model, data = test_df)$predictions
        rf_r2 <- 1 - sum((test_df[[vaisala_col]] - rf_preds)^2) / ss_tot
      } else {
        rf_r2 <- NA_real_
      }
      
      # LM (Always fit for Ozone and as fallback for PM)
      lm_model <- lm(rf_formula, data = train_df)
      lm_preds <- predict(lm_model, newdata = test_df)
      lm_r2 <- 1 - sum((test_df[[vaisala_col]] - lm_preds)^2) / ss_tot
    } else {
      rf_r2 <- lm_r2 <- NA_real_
    }
    
    # Selection logic (optimized for speed and user request)
    if (param == "o3_ppm") {
      best_model <- lm_model
      test_r2 <- lm_r2
      model_type <- "lm"
      cat(sprintf("  Selection: Forced Linear Model for Ozone (Test R²: %.3f)\n", lm_r2))
    } else {
      # For PM: Keep Random Forest as the primary approach
      # Use RF if R2 > 0. If RF is very poor (R2 < 0) but LM is better, use LM.
      # Otherwise fallback to bias.
      use_rf <- (!is.na(rf_r2) && rf_r2 >= 0)
      use_lm <- (!use_rf && !is.na(lm_r2) && lm_r2 >= 0)
      
      if (use_rf) {
        best_model <- rf_model
        test_r2 <- rf_r2
        model_type <- "rf"
        cat(sprintf("  Selection: Using Random Forest (Test R²: %.3f)\n", rf_r2))
      } else if (use_lm) {
        best_model <- lm_model
        test_r2 <- lm_r2
        model_type <- "lm"
        cat(sprintf("  Selection: Using Linear Model (Test R²: %.3f)\n", lm_r2))
      } else {
        # FALLBACK: Bias-correction (Mean of training Vaisala data)
        bias_val <- mean(train_df[[vaisala_col]], na.rm = TRUE)
        best_model <- list(bias_val = bias_val)
        test_r2 <- 0 
        model_type <- "bias_only"
        cat(sprintf("  Selection: Falling back to BIAS-CORRECTION (Mean shift) due to poor performance (RF R²: %.3f, LM R²: %.3f)\n", 
                    rf_r2, lm_r2))
      }
    }

    # RMSE on test set
    best_preds <- if(model_type == "rf") rf_preds else if(model_type == "lm") lm_preds else rep(best_model$bias_val, nrow(test_df))
    rmse_val <- sqrt(mean((test_df[[vaisala_col]] - best_preds)^2))

    # Variable importance
    vi <- if(model_type == "rf") rf_model$variable.importance else NULL

    general_models[[param]] <- list(
      parameter  = param,
      model      = best_model,
      model_type = model_type,
      bias_value = if(model_type == "bias_only") best_model$bias_val else NULL,
      r_squared  = test_r2,
      rmse       = rmse_val,
      n_train    = n_train,
      n_test     = nrow(test_df),
      n_points   = nrow(model_data),
      importance = vi,
      equation   = paste0(toupper(model_type), " model: ", param,
                          " ~ Arduino + RH + Temp | test R² = ", round(test_r2, 3))
    )

    cat(sprintf("\n=== General %s Calibration Model for %s ===\n", toupper(model_type), param))
    cat("Test-set R²:", round(test_r2, 4), "\n")
    cat("Test-set RMSE:", round(rmse_val, 2), "\n")
    cat("Train / test rows:", n_train, "/", nrow(test_df), "\n")
    if(!is.null(vi)) {
      cat("Variable importance:\n")
      print(round(vi, 2))
    }
    cat("\n")
  }

  return(general_models)
}

# Create models
general_models <- create_general_calibration_models(all_data)
```

# Site-Specific Calibration Data Loading
```{r}
# Load site-specific calibration data from Cal_Stations directory
# Each site has its own folder containing:
#   - Site sensor data CSV file
#   - VaisalaTable.csv (reference data)

cal_stations_dir <- file.path(dirname(data_dir), "Cal_Stations")

# Function to read CSV files (handles different formats)
read_calibration_csv <- function(file_path, site_name) {
  if(!file.exists(file_path)) {
    cat("File does not exist:", basename(file_path), "\n")
    return(NULL)
  }
  
  # Try reading with different approaches
  encodings <- c("UTF-8", "latin1", "ISO-8859-1", "windows-1252")
  
  for(enc in encodings) {
    tryCatch({
      # First, try reading without skip (in case it's a standard CSV)
      data <- read.csv(file_path, header=TRUE, 
                      fileEncoding=enc, stringsAsFactors=FALSE,
                      check.names=FALSE, quote="\"", comment.char="",
                      as.is=TRUE)
      
      if(nrow(data) > 0 && ncol(data) > 0) {
        cat("Successfully loaded", basename(file_path), "with encoding:", enc, "\n")
        cat("  Columns:", paste(colnames(data), collapse=", "), "\n")
        cat("  Rows:", nrow(data), "\n")
        return(data)
      }
    }, error = function(e) {
      return(NULL)
    })
  }
  
  # Try with skip=3 (in case it has header rows like the .dat files)
  for(enc in encodings) {
    tryCatch({
      data <- read.csv(file_path, header=TRUE, skip=3,
                      fileEncoding=enc, stringsAsFactors=FALSE,
                      check.names=FALSE, quote="\"", comment.char="",
                      as.is=TRUE)
      
      if(nrow(data) > 0 && ncol(data) > 0) {
        cat("Successfully loaded", basename(file_path), "with encoding:", enc, "(skip=3)\n")
        cat("  Columns:", paste(colnames(data), collapse=", "), "\n")
        cat("  Rows:", nrow(data), "\n")
        return(data)
      }
    }, error = function(e) {
      return(NULL)
    })
  }
  
  cat("Failed to load", basename(file_path), "\n")
  return(NULL)
}

# Scan Cal_Stations directory for site folders
site_calibration_data <- list()

if(dir.exists(cal_stations_dir)) {
  cat("Scanning Cal_Stations directory for site-specific calibration data...\n")
  
  # Get all subdirectories (site folders)
  site_folders <- list.dirs(cal_stations_dir, full.names = TRUE, recursive = FALSE)
  
  cat("Found", length(site_folders), "site folder(s)\n\n")
  
  for(site_folder in site_folders) {
    site_name <- basename(site_folder)
    cat("Processing site:", site_name, "\n")
    
    # Find files in this folder
    files_in_folder <- list.files(site_folder, full.names = TRUE)
    
    # Look for Vaisala file and site sensor file
    vaisala_file <- NULL
    sensor_file <- NULL
    
    for(f in files_in_folder) {
      f_basename <- basename(f)
      if(tolower(f_basename) == "vaisalatable.csv") {
        vaisala_file <- f
      } else if(grepl("\\.csv$", f_basename, ignore.case = TRUE)) {
        # This should be the site sensor data file
        sensor_file <- f
      }
    }
    
    if(is.null(vaisala_file)) {
      cat("  Warning: No VaisalaTable.csv found in", site_name, "\n")
      next
    }
    
    if(is.null(sensor_file)) {
      cat("  Warning: No sensor data CSV found in", site_name, "\n")
      next
    }
    
    # Load both files
    cat("  Loading Vaisala data:", basename(vaisala_file), "\n")
    vaisala_data <- read_calibration_csv(vaisala_file, paste0(site_name, "_Vaisala"))
    
    cat("  Loading sensor data:", basename(sensor_file), "\n")
    sensor_data <- read_calibration_csv(sensor_file, site_name)
    
    if(!is.null(vaisala_data) && !is.null(sensor_data)) {
      site_calibration_data[[site_name]] <- list(
        site_name = site_name,
        vaisala_data = vaisala_data,
        sensor_data = sensor_data
      )
      cat("  Successfully loaded data for", site_name, "\n")
    } else {
      cat("  Failed to load complete data for", site_name, "\n")
    }
    cat("\n")
  }
  
  cat("Summary: Loaded calibration data for", length(site_calibration_data), "site(s)\n")
  cat("Sites:", paste(names(site_calibration_data), collapse = ", "), "\n\n")
} else {
  cat("Cal_Stations directory not found:", cal_stations_dir, "\n")
  cat("Falling back to old file locations...\n")
  
  # Fallback to old file locations for backward compatibility
  vaisala_new_file <- file.path(data_dir, "VaisalaTable.csv")
  strength2love_file <- file.path(data_dir, "Strength2LoveFarm_SERCCal.csv")
  baybrook_file <- file.path(data_dir, "BayBrookElementary_SERCCal.csv")
  
  if(file.exists(vaisala_new_file) && file.exists(strength2love_file)) {
    vaisala_data <- read_calibration_csv(vaisala_new_file, "Vaisala_New")
    sensor_data <- read_calibration_csv(strength2love_file, "Strength2LoveFarm")
    if(!is.null(vaisala_data) && !is.null(sensor_data)) {
      site_calibration_data[["Strength2LoveFarm"]] <- list(
        site_name = "Strength2LoveFarm",
        vaisala_data = vaisala_data,
        sensor_data = sensor_data
      )
    }
  }
  
  if(file.exists(vaisala_new_file) && file.exists(baybrook_file)) {
    vaisala_data <- read_calibration_csv(vaisala_new_file, "Vaisala_New")
    sensor_data <- read_calibration_csv(baybrook_file, "BayBrookElementary")
    if(!is.null(vaisala_data) && !is.null(sensor_data)) {
      site_calibration_data[["BayBrookElementary"]] <- list(
        site_name = "BayBrookElementary",
        vaisala_data = vaisala_data,
        sensor_data = sensor_data
      )
    }
  }
}
```

# Site-Specific Data Processing
```{r}
# Function to standardize column names and extract relevant measurements
standardize_sensor_data <- function(data, site_name) {
  if(is.null(data) || nrow(data) == 0) {
    return(NULL)
  }
  
  # Convert column names to lowercase for easier matching
  colnames_lower <- tolower(colnames(data))
  data_cols <- colnames(data)
  
  # Helper: find a column by exact name match (case-insensitive), returns first match
  find_column <- function(candidates) {
    for(name in candidates) {
      idx <- which(colnames_lower == tolower(name))
      if(length(idx) > 0) return(data_cols[idx[1]])
    }
    return(NULL)
  }
  
  # Find timestamp column
  timestamp_col <- find_column(c("TIMESTAMP", "timestamp", "datetime", "date", "time", "date_time"))
  
  if(is.null(timestamp_col)) {
    cat("Warning: Could not find timestamp column in", site_name, "\n")
    cat("Available columns:", paste(colnames(data), collapse=", "), "\n")
    return(NULL)
  }
  
  # Extract and convert timestamp
  result <- data.frame(timestamp = data[[timestamp_col]], stringsAsFactors = FALSE)
  
  # Parse timestamps — convert incoming ET/EST to UTC
  result$timestamp <- parse_to_utc(result$timestamp)
  
  # --- Column matching ---
  # Uses exact name matching (case-insensitive) with explicit candidate lists.
  # Arduino data logger naming convention:
  #   ard_pm10_1  = PM1.0  (1.0 μm × 10 = 10)  sensor 1
  #   ard_pm25_1  = PM2.5  (2.5 μm × 10 = 25)  sensor 1
  #   ard_pm100_1 = PM10   (10  μm × 10 = 100) sensor 1
  # Some units use ard2_ prefix instead of ard_.
  
  # O3
  o3_col <- find_column(c(
    "O3", "o3_ppb", "ozone_ppb", "ozone",
    "ard_o3", "ard2_o3",                    # Cal_Stations Arduino names
    "ard_o3_ppb"                             # renamed general calibration name
  ))
  
  # PM1.0  (note: ard_pm10_1 = PM1.0, NOT PM10)
  pm1_col <- find_column(c(
    "PM1", "pm1_ugm3",
    "ard_pm10_1", "ard2_pm10_1",            # Cal_Stations Arduino names (PM1.0)
    "ard_pm1_1_ugm3"                         # renamed general calibration name
  ))
  
  # PM2.5
  pm25_col <- find_column(c(
    "PM2_5", "pm25_ugm3", "pm2.5",
    "ard_pm25_1", "ard2_pm25_1",            # Cal_Stations Arduino names
    "ard_pm25_1_ugm3"                        # renamed general calibration name
  ))
  
  # PM10  (note: ard_pm100_1 = PM10, NOT PM100)
  pm10_col <- find_column(c(
    "PM10", "pm10_ugm3",
    "ard_pm100_1", "ard2_pm100_1",          # Cal_Stations Arduino names (PM10)
    "ard_pm10_1_ugm3"                        # renamed general calibration name
  ))
  
  # Temperature
  temp_col <- find_column(c(
    "air_temp",                              # Vaisala
    "ard_temp", "ard2_temp",                 # Cal_Stations Arduino names
    "ard_temp_c",                            # renamed general calibration name
    "temp_c", "temperature", "bme_temp"
  ))
  
  # Humidity
  humidity_col <- find_column(c(
    "humidity",                              # Vaisala
    "ard_rh", "ard2_rh",                    # Cal_Stations Arduino names
    "ard_humidity_perc_rh",                  # renamed general calibration name
    "humidity_perc_rh", "rh", "bme_humidity"
  ))
  
  # Extract measurements
  if(!is.null(o3_col)) result$o3_ppm <- as.numeric(data[[o3_col]]) / 1000
  if(!is.null(pm1_col)) result$pm1_ugm3 <- as.numeric(data[[pm1_col]])
  if(!is.null(pm25_col)) result$pm25_ugm3 <- as.numeric(data[[pm25_col]])
  if(!is.null(pm10_col)) result$pm10_ugm3 <- as.numeric(data[[pm10_col]])
  if(!is.null(temp_col)) result$temp_c <- as.numeric(data[[temp_col]])
  if(!is.null(humidity_col)) result$humidity_perc_rh <- as.numeric(data[[humidity_col]])
  
  result$site <- site_name
  
  cat("\nStandardized data for", site_name, ":\n")
  cat("  Timestamp column:", timestamp_col, "\n")
  cat("  O3 column:", ifelse(is.null(o3_col), "NOT FOUND", o3_col), "\n")
  cat("  PM1 column:", ifelse(is.null(pm1_col), "NOT FOUND", pm1_col), "\n")
  cat("  PM2.5 column:", ifelse(is.null(pm25_col), "NOT FOUND", pm25_col), "\n")
  cat("  PM10 column:", ifelse(is.null(pm10_col), "NOT FOUND", pm10_col), "\n")
  cat("  Temp column:", ifelse(is.null(temp_col), "NOT FOUND", temp_col), "\n")
  cat("  Humidity column:", ifelse(is.null(humidity_col), "NOT FOUND", humidity_col), "\n")
  cat("  Rows:", nrow(result), "\n")
  
  return(result)
}

# Standardize all site-specific data files
site_calibration_std <- list()

for(site_name in names(site_calibration_data)) {
  site_info <- site_calibration_data[[site_name]]
  
  cat("\nStandardizing data for", site_name, "...\n")
  
  # Standardize sensor data
  sensor_std <- standardize_sensor_data(site_info$sensor_data, site_name)
  
  # Standardize Vaisala data
  vaisala_std <- standardize_sensor_data(site_info$vaisala_data, paste0(site_name, "_Vaisala"))
  
  if(!is.null(sensor_std) && !is.null(vaisala_std)) {
    site_calibration_std[[site_name]] <- list(
      site_name = site_name,
      sensor_data = sensor_std,
      vaisala_data = vaisala_std
    )
    cat("Successfully standardized data for", site_name, "\n")
  } else {
    cat("Failed to standardize data for", site_name, "\n")
  }
}

cat("\nSummary: Standardized data for", length(site_calibration_std), "site(s)\n")
cat("Sites:", paste(names(site_calibration_std), collapse = ", "), "\n")
```

# Site-Specific Calibration Model Creation
```{r site-specific-models}
# Creates a random-forest calibration model for a specific site.
# Predictors: sensor reading + temperature + relative humidity
# 80/20 temporal train/test split; reports test-set R² and variable importance.
# Also stores combined_data (hourly colocation) for the second-stage linear correction.
create_site_calibration_models <- function(sensor_data, vaisala_data, site_name) {
  if(is.null(sensor_data) || is.null(vaisala_data)) {
    cat("Missing data for", site_name, "\n")
    return(NULL)
  }

  # Ensure timestamps are POSIXct in UTC
  sensor_data$timestamp <- parse_to_utc(sensor_data$timestamp)
  vaisala_data$timestamp <- parse_to_utc(vaisala_data$timestamp)

  # Round timestamps to nearest hour for matching
  sensor_data$timestamp_hour <- round(sensor_data$timestamp, units = "hours")
  vaisala_data$timestamp_hour <- round(vaisala_data$timestamp, units = "hours")

  # Determine which measurement columns are available
  all_measure_cols <- c("o3_ppm", "pm1_ugm3", "pm25_ugm3", "pm10_ugm3", "temp_c", "humidity_perc_rh")
  sensor_measure_cols <- intersect(all_measure_cols, colnames(sensor_data))
  vaisala_measure_cols <- intersect(all_measure_cols, colnames(vaisala_data))

  cat("  Sensor columns available:", paste(sensor_measure_cols, collapse = ", "), "\n")
  cat("  Vaisala columns available:", paste(vaisala_measure_cols, collapse = ", "), "\n")

  if(length(sensor_measure_cols) == 0 || length(vaisala_measure_cols) == 0) {
    cat("No overlapping measurement columns for", site_name, "\n")
    return(NULL)
  }

  # Aggregate to hourly means
  sensor_hourly <- sensor_data %>%
    group_by(timestamp_hour) %>%
    summarise(across(all_of(sensor_measure_cols), ~mean(.x, na.rm = TRUE)), .groups = 'drop')

  vaisala_hourly <- vaisala_data %>%
    group_by(timestamp_hour) %>%
    summarise(across(all_of(vaisala_measure_cols), ~mean(.x, na.rm = TRUE)), .groups = 'drop')

  # Merge by timestamp
  combined_data <- inner_join(
    sensor_hourly,
    vaisala_hourly,
    by = "timestamp_hour",
    suffix = c("_sensor", "_vaisala")
  )

  n_matched <- nrow(combined_data)

  # --- Step 2: 72-hour warning ---
  if(site_name %in% calibrated_sensors && n_matched < 72) {
    warning(sprintf(
      "Calibrated sensor '%s' has only %d hourly observations in its colocation period (< 72 recommended). Site-specific correction may be unreliable.",
      site_name, n_matched
    ))
  }

  if(n_matched < min_hours_for_calibration) {
    cat("Insufficient overlapping data for", site_name,
        "(", n_matched, "hourly points; need", min_hours_for_calibration, ")\n")
    return(NULL)
  }

  cat("\n=== Creating RF calibration models for", site_name, "===\n")
  cat("Matched", n_matched, "hourly data points\n")

  # Model Rationale: 
  # - Ozone: Forced MLR is used to handle low-variance "flat signals."
  # - PM: Random Forest captures non-linear RH/Temp interactions.
  # - Efficiency: Multi-core processing enabled; redundant fits skipped.
  
  target_params <- c("o3_ppm", "pm1_ugm3", "pm25_ugm3", "pm10_ugm3")
  models <- list()

  for(param in target_params) {
    sensor_col    <- paste0(param, "_sensor")
    vaisala_col   <- paste0(param, "_vaisala")

    if(!(sensor_col %in% colnames(combined_data)) ||
       !(vaisala_col %in% colnames(combined_data))) {
      cat("Missing columns for", param, "\n")
      next
    }

    has_env <- "temp_c_sensor" %in% colnames(combined_data) &&
               "humidity_perc_rh_sensor" %in% colnames(combined_data)

    filter_expr <- if(has_env) {
      combined_data %>%
        filter(!is.na(.data[[sensor_col]]), !is.na(.data[[vaisala_col]]),
               is.finite(.data[[sensor_col]]), is.finite(.data[[vaisala_col]]),
               !is.na(temp_c_sensor), !is.na(humidity_perc_rh_sensor),
               is.finite(temp_c_sensor), is.finite(humidity_perc_rh_sensor))
    } else {
      combined_data %>%
        filter(!is.na(.data[[sensor_col]]), !is.na(.data[[vaisala_col]]),
               is.finite(.data[[sensor_col]]), is.finite(.data[[vaisala_col]]))
    }
    model_data <- filter_expr

    if(nrow(model_data) < min_hours_for_calibration) next

    # 80/20 random stratified split (replacing temporal split)
    set.seed(42)
    n_train <- floor(0.8 * nrow(model_data))
    train_indices <- sample(seq_len(nrow(model_data)), n_train)
    train_df <- model_data[train_indices, ]
    test_df  <- model_data[-train_indices, ]

    # --- Diagnostics (per user request) ---
    cat(sprintf("\n--- Diagnostics for %s (%s) ---\n", param, site_name))
    cat("  Training date range:", as.character(min(as.Date(train_df$timestamp_hour))), 
        "to", as.character(max(as.Date(train_df$timestamp_hour))), "\n")
    cat("  Unique dates in training:", length(unique(as.Date(train_df$timestamp_hour))), "\n")
    
    # Correlation check (verify alignment)
    cor_val <- cor(train_df[[sensor_col]], train_df[[vaisala_col]], use = "complete.obs")
    cat(sprintf("  Correlation (%s vs %s): %.3f\n", sensor_col, vaisala_col, cor_val))

    # Build predictor data frames
    if(has_env) {
      rf_formula <- as.formula(paste0(
        vaisala_col, " ~ ", sensor_col, " + temp_c_sensor + humidity_perc_rh_sensor"
      ))
    } else {
      rf_formula <- as.formula(paste0(vaisala_col, " ~ ", sensor_col))
    }

    # Fit models (optimized for speed: skip RF for Ozone)
    ss_tot <- sum((test_df[[vaisala_col]] - mean(test_df[[vaisala_col]]))^2)
    rf_preds <- NULL
    
    if (ss_tot > 0) {
    if (param != "o3_ppm") {
        rf_model <- ranger(
          formula    = rf_formula,
          data       = train_df,
          importance = "impurity",
          num.trees  = 500,
          seed       = 42,
          num.threads = parallel::detectCores() # Speed optimization
        )
        rf_preds <- predict(rf_model, data = test_df)$predictions
        rf_r2 <- 1 - sum((test_df[[vaisala_col]] - rf_preds)^2) / ss_tot
      } else {
        rf_r2 <- NA_real_
      }
      
      # LM (Always fit for Ozone and as fallback for PM)
      lm_model <- lm(rf_formula, data = train_df)
      lm_preds <- predict(lm_model, newdata = test_df)
      lm_r2 <- 1 - sum((test_df[[vaisala_col]] - lm_preds)^2) / ss_tot
    } else {
      rf_r2 <- lm_r2 <- NA_real_
    }
    
    # Selection logic (optimized for speed and user request)
    if (param == "o3_ppm") {
      best_model <- lm_model
      test_r2 <- lm_r2
      model_type <- "lm"
      cat(sprintf("  Selection: Forced Linear Model for Ozone (Test R²: %.3f)\n", lm_r2))
    } else {
      # For PM: Keep Random Forest as the primary approach
      use_rf <- (!is.na(rf_r2) && rf_r2 >= 0)
      use_lm <- (!use_rf && !is.na(lm_r2) && lm_r2 >= 0)
      
      if (use_rf) {
        best_model <- rf_model
        test_r2 <- rf_r2
        model_type <- "rf"
        cat(sprintf("  Selection: Using Random Forest (Test R²: %.3f)\n", rf_r2))
      } else if (use_lm) {
        best_model <- lm_model
        test_r2 <- lm_r2
        model_type <- "lm"
        cat(sprintf("  Selection: Using Linear Model (Test R²: %.3f)\n", lm_r2))
      } else {
        # FALLBACK: Bias-correction
        bias_val <- mean(train_df[[vaisala_col]], na.rm = TRUE)
        best_model <- list(bias_val = bias_val)
        test_r2 <- 0 
        model_type <- "bias_only"
        cat(sprintf("  Selection: Falling back to BIAS-CORRECTION (Mean shift) due to poor performance (RF R²: %.3f, LM R²: %.3f)\n", 
                    rf_r2, lm_r2))
      }
    }

    # RMSE on test set
    best_preds <- if(model_type == "rf") rf_preds else if(model_type == "lm") lm_preds else rep(best_model$bias_val, nrow(test_df))
    rmse_val <- sqrt(mean((test_df[[vaisala_col]] - best_preds)^2))

    # Variable importance
    vi <- if(model_type == "rf") rf_model$variable.importance else NULL

    models[[param]] <- list(
      parameter   = param,
      model       = best_model,
      model_type  = model_type,
      bias_value  = if(model_type == "bias_only") best_model$bias_val else NULL,
      r_squared   = test_r2,          # test-set R²
      rmse        = rmse_val,
      n_train     = n_train,
      n_test      = nrow(test_df),
      n_points    = nrow(model_data),
      importance  = vi,
      site        = site_name,
      train_start = min(train_df$timestamp_hour),
      train_end   = max(train_df$timestamp_hour),
      equation     = paste0(toupper(model_type), " model: ", param,
                            " ~ sensor + temp + RH | test R² = ",
                            round(test_r2, 3))
    )

    cat(sprintf("\n=== %s Calibration for %s (%s) ===\n", toupper(model_type), param, site_name))
    cat("Test-set R²:", round(test_r2, 4), "\n")
    cat("Test-set RMSE:", round(rmse_val, 2), "\n")
    cat("Train / test rows:", n_train, "/", nrow(test_df), "\n")
    if(!is.null(vi)) {
      cat("Variable importance:\n")
      print(round(vi, 2))
    }
  }

  if(length(models) == 0) {
    cat("No calibration models could be created for", site_name, "\n")
    return(NULL)
  }

  return(list(models = models, combined_data = combined_data))
}

# Create calibration models for each site
site_calibration_models <- list()

for(site_name in names(site_calibration_std)) {
  site_info <- site_calibration_std[[site_name]]

  cat("\n=== Creating calibration models for", site_name, "===\n")

  models_result <- create_site_calibration_models(
    site_info$sensor_data,
    site_info$vaisala_data,
    site_name
  )

  if(!is.null(models_result) && length(models_result$models) > 0) {
    site_calibration_models[[site_name]] <- models_result
    cat("Successfully created models for", site_name,
        "(", paste(names(models_result$models), collapse = ", "), ")\n")
  } else {
    cat("Failed to create models for", site_name, "\n")
  }
}

cat("\nSummary: Created calibration models for", length(site_calibration_models), "site(s)\n")
cat("Sites with models:", paste(names(site_calibration_models), collapse = ", "), "\n")
```

# Site-Specific Calibration Visualizations
```{r site-viz, fig.width=10, fig.height=5}
# Function to create calibration plots for each site
create_site_calibration_plots <- function(models_result, site_name) {
  if(is.null(models_result) || is.null(models_result$models)) {
    cat("No models available for", site_name, "\n")
    return(NULL)
  }
  
  models <- models_result$models
  combined_data <- models_result$combined_data
  
  if(is.null(combined_data) || nrow(combined_data) == 0) {
    return(NULL)
  }
  
  parameters <- list(
    "o3_ppm" = "Ozone (ppm)",
    "pm1_ugm3" = "PM1 (μg/m³)",
    "pm25_ugm3" = "PM2.5 (μg/m³)",
    "pm10_ugm3" = "PM10 (μg/m³)"
  )
  
  for(param in names(parameters)) {
    if(!param %in% names(models)) next
    
    param_label <- parameters[[param]]
    sensor_col <- paste0(param, "_sensor")
    vaisala_col <- paste0(param, "_vaisala")
    
    if(!(sensor_col %in% colnames(combined_data)) || 
       !(vaisala_col %in% colnames(combined_data))) next
    
    plot_data <- combined_data %>%
      filter(
        !is.na(.data[[sensor_col]]),
        !is.na(.data[[vaisala_col]]),
        is.finite(.data[[sensor_col]]),
        is.finite(.data[[vaisala_col]])
      )
    
    if(nrow(plot_data) < 3) next
    
    model_info <- models[[param]]
    
    # Scatter plot with regression line
    p1 <- ggplot(plot_data, aes(x = .data[[sensor_col]], y = .data[[vaisala_col]])) +
      geom_point(alpha = 0.6, color = "steelblue") +
      geom_smooth(method = "lm", se = TRUE, color = "red", linetype = "dashed") +
      geom_abline(intercept = 0, slope = 1, color = "gray", linetype = "dotted") +
      labs(
        title = paste(site_name, "-", param_label, "Calibration"),
        subtitle = paste0("R² = ", round(model_info$r_squared, 3), 
                          ", RMSE = ", round(model_info$rmse, 2)),
        x = paste("Sensor", param_label),
        y = paste("Vaisala Reference", param_label)
      ) +
      theme_minimal()
    
    print(p1)
    
    # Time series comparison
    p2 <- ggplot(plot_data, aes(x = timestamp_hour)) +
      geom_line(aes(y = .data[[sensor_col]], color = "Sensor"), alpha = 0.7) +
      geom_line(aes(y = .data[[vaisala_col]], color = "Vaisala Reference"), alpha = 0.7) +
      labs(
        title = paste(site_name, "-", param_label, "Time Series"),
        x = "Timestamp",
        y = param_label,
        color = "Source"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_color_manual(values = c("Sensor" = "blue", "Vaisala Reference" = "green"))
    
    print(p2)
  }
}

# Create plots for each site
for(site_name in names(site_calibration_models)) {
  cat("\n=== Creating plots for", site_name, "===\n")
  create_site_calibration_plots(site_calibration_models[[site_name]], site_name)
}
```

# Site-Specific Calibration Summary
```{r site-summary}
# Print summary of all calibration equations
cat("\n\n")
cat(paste0(rep("=", 80), collapse=""), "\n")
cat("SITE-SPECIFIC CALIBRATION EQUATIONS SUMMARY\n")
cat(paste0(rep("=", 80), collapse=""), "\n\n")

for(site_name in names(site_calibration_models)) {
  models_result <- site_calibration_models[[site_name]]
  
  if(!is.null(models_result) && !is.null(models_result$models)) {
    cat(toupper(site_name), "CALIBRATION EQUATIONS:\n")
    cat(paste0(rep("-", 80), collapse=""), "\n")
    
    for(param in names(models_result$models)) {
      model_info <- models_result$models[[param]]
      cat("\n", toupper(param), ":\n")
      cat("  ", model_info$equation, "\n")
      cat("  R² =", round(model_info$r_squared, 4), 
          ", RMSE =", round(model_info$rmse, 2),
          ", N =", model_info$n_points, "\n")
    }
    cat("\n")
  }
}

cat(paste0(rep("=", 80), collapse=""), "\n\n")
```

# -------------------------------------------------------------------------
# CALIBRATION DEPLOYMENT SUMMARY
# -------------------------------------------------------------------------
```{r deployment-summary}
# Summary table of applied models and date ranges.
sites_with_calibration <- names(site_calibration_models)

# Create a detailed deployment summary table
deployment_rows <- list()

# Detailed descriptions for General Stage 1
gen_types <- paste(sapply(names(general_models), function(p) {
  paste0(toupper(p), ": ", toupper(general_models[[p]]$model_type))
}), collapse = " | ")

gen_eqs <- paste(sapply(names(general_models), function(p) {
  m <- general_models[[p]]
  if(m$model_type == "lm") {
    co <- coef(m$model)
    sprintf("%s: %.2f + %.2f*x", toupper(p), co[1], co[2])
  } else {
    paste0(toupper(p), ": R²=", round(m$r_squared, 2))
  }
}), collapse = " | ")

for(site in sites_with_calibration) {
  # Get a representative model to check the training range
  site_models <- site_calibration_models[[site]]$models
  rep_model <- site_models[[1]]
  
  # Before cutoff
  deployment_rows[[length(deployment_rows) + 1]] <- data.frame(
    Site = site,
    Application_Period = paste0("Before ", calibration_cutoff_date),
    Applied_Model = "General Stage 1",
    Model_Type = gen_types,
    Equations = gen_eqs,
    Basis = "Lab co-location",
    stringsAsFactors = FALSE
  )
  
  # After cutoff: Site-specific details
  site_types <- paste(sapply(site_models, function(m) {
    paste0(toupper(m$parameter), ": ", toupper(m$model_type))
  }), collapse = " | ")
  
  site_eqs <- paste(sapply(site_models, function(m) {
    if(m$model_type == "lm") {
      co <- coef(m$model)
      sprintf("%s: %.2f + %.2f*x", toupper(m$parameter), co[1], co[2])
    } else {
      paste0(toupper(m$parameter), ": R²=", round(m$r_squared, 2))
    }
  }), collapse = " | ")
  
  deployment_rows[[length(deployment_rows) + 1]] <- data.frame(
    Site = site,
    Application_Period = paste0("On/After ", calibration_cutoff_date),
    Applied_Model = "Site-Specific Stage 2",
    Model_Type = site_types,
    Equations = site_eqs,
    Basis = sprintf("Field co-location (%s to %s)", 
                    as.Date(rep_model$train_start), 
                    as.Date(rep_model$train_end)),
    stringsAsFactors = FALSE
  )
}

# Add default for other sites
deployment_rows[[length(deployment_rows) + 1]] <- data.frame(
  Site = "All Other Sites",
  Application_Period = "All Dates",
  Applied_Model = "General Stage 1",
  Model_Type = gen_types,
  Equations = gen_eqs,
  Basis = "Lab co-location",
  stringsAsFactors = FALSE
)

calibration_deployment_summary <- bind_rows(deployment_rows)

cat("\n")
cat(paste0(rep("=", 80), collapse=""), "\n")
cat("DETAILED CALIBRATION DEPLOYMENT BY SITE AND DATE\n")
cat(paste0(rep("=", 80), collapse=""), "\n\n")
print(calibration_deployment_summary)
cat("\n")
cat(paste0(rep("=", 80), collapse=""), "\n\n")
```

# Calibration Usage Guide
```{r}
# Note: calibration_cutoff_date is defined in the previous section

cat("\n")
cat(paste0(rep("=", 80), collapse=""), "\n")
cat("CALIBRATION EQUATIONS USAGE GUIDE\n")
cat(paste0(rep("=", 80), collapse=""), "\n\n")
cat("IMPORTANT: Date-based calibration selection\n")
cat("  Calibration cutoff date:", as.character(calibration_cutoff_date), "\n\n")

if(length(sites_with_calibration) > 0) {
  cat("SITE-SPECIFIC CALIBRATIONS:\n")
  cat("  Sites with site-specific calibrations:", 
      paste(sites_with_calibration, collapse = ", "), "\n")
  cat("  - ONLY use for data collected ON OR AFTER", as.character(calibration_cutoff_date), "\n")
  cat("  - These calibrations are based on co-located Vaisala data\n")
  cat("  - For data BEFORE", as.character(calibration_cutoff_date), ", use GENERAL calibration\n\n")
} else {
  cat("SITE-SPECIFIC CALIBRATIONS:\n")
  cat("  No site-specific calibrations available.\n\n")
}

cat("GENERAL CALIBRATION (All sites):\n")
cat("  - Use for ALL data collected BEFORE", as.character(calibration_cutoff_date), "\n")
if(length(sites_with_calibration) > 0) {
  cat("  - Use for sites without site-specific calibration at ALL times\n")
  cat("  - Use for", paste(sites_with_calibration, collapse = ", "), 
      "BEFORE", as.character(calibration_cutoff_date), "\n")
}
cat("  - Based on the original Arduino vs Vaisala comparison\n\n")

cat("When applying corrections:\n")
cat("  1. Check the date of the measurement\n")
if(length(sites_with_calibration) > 0) {
  cat("  2. If date >= ", as.character(calibration_cutoff_date), 
      " AND site is one of:", paste(sites_with_calibration, collapse = ", "), "\n")
  cat("     -> Use site-specific calibration\n")
  cat("  3. Otherwise:\n")
} else {
  cat("  2. Always:\n")
}
cat("     -> Use general calibration\n")
cat(paste0(rep("=", 80), collapse=""), "\n\n")
```

# -------------------------------------------------------------------------
# CALIBRATION APPLICATION (PRODUCTION)
# -------------------------------------------------------------------------
# Stage 1: General RF/MLR (All sensors, all dates)
# Stage 2: Site-Specific Linear Refinement (Calibrated sites, date >= cutoff)
# -------------------------------------------------------------------------
```{r apply-calibration}
apply_calibration_by_site <- function(sensor_data, site_name,
                                     general_models = NULL,
                                     site_calibration_models = NULL,
                                     colocation_data_list = NULL,
                                     timestamp_col = "timestamp") {
  if(is.null(sensor_data) || nrow(sensor_data) == 0) {
    cat("No sensor data provided for", site_name, "\n")
    return(NULL)
  }

  if(!timestamp_col %in% colnames(sensor_data)) {
    cat("Timestamp column", timestamp_col, "not found in data\n")
    return(NULL)
  }

  sensor_data$date <- as.Date(sensor_data[[timestamp_col]])

  # ── Stage 1: Apply general RF to ALL rows ──────────────────────────────
  corrected_data <- apply_rf_calibration(sensor_data, general_models, site_name)
  if(is.null(corrected_data)) corrected_data <- sensor_data

  # Set default method for every row
  corrected_data$correction_method <- "rf_general"

  # ── Stage 2: Site-specific linear correction (calibrated sensors only) ──
  # Only for sensors in calibrated_sensors AND date >= calibration_cutoff_date
  is_calibrated <- site_name %in% calibrated_sensors

  if(is_calibrated) {
    # Find colocation data for this site
    cal_data <- NULL
    if(!is.null(colocation_data_list) && site_name %in% names(colocation_data_list)) {
      cal_data <- colocation_data_list[[site_name]]
    } else if(!is.null(site_calibration_models) && site_name %in% names(site_calibration_models)) {
      cal_data <- site_calibration_models[[site_name]]$combined_data
    }

    post_rows <- corrected_data$date >= calibration_cutoff_date

    if(any(post_rows) && !is.null(cal_data)) {
      cat("\n  Stage 2: fitting site-specific linear correction for", site_name, "\n")
      target_params <- c("o3_ppb", "pm1_ugm3", "pm25_ugm3", "pm10_ugm3")

      for(param in target_params) {
        rf_col  <- paste0("corrected_", param)      # RF stage-1 corrected
        lab_col <- paste0(param, "_vaisala")         # lab (Vaisala) reference

        if(!(rf_col %in% colnames(corrected_data))) {
          cat("    No RF-corrected column for", param, "— skipping\n")
          next
        }

        # Build calibration data: need rf_col in corrected form for colocation period
        # Re-run stage-1 RF on the colocation data to get RF-corrected values there
        col_rf <- NULL
        if(!is.null(general_models) && param %in% names(general_models)) {
          # colocation sensor column
          sensor_param_col <- paste0(param, "_sensor")
          if(sensor_param_col %in% colnames(cal_data)) {
            pred_df <- build_rf_pred_df(cal_data, param, type = "site")
            if(!is.null(pred_df)) {
              # General model expects _arduino names even when applying to site data
              colnames(pred_df) <- c(paste0(param, "_arduino"), 
                                     "humidity_perc_rh_arduino", 
                                     "temp_c_arduino")
              # Use helper to apply General Model
              col_rf <- predict_best_model(general_models[[param]], pred_df)
            }
          } else {
            # fall back: try the arduino-style naming used in all_data
            arduino_col <- paste0(param, "_arduino")
            if(arduino_col %in% colnames(cal_data)) {
              pred_df <- build_rf_pred_df(cal_data, param, type = "arduino")
              if(!is.null(pred_df)) {
                # Use helper to apply General Model
                col_rf <- predict_best_model(general_models[[param]], pred_df)
              }
            }
          }
        }

        if(is.null(col_rf) || !(lab_col %in% colnames(cal_data))) {
          cat("    Cannot build stage-2 data for", param, "— skipping\n")
          next
        }

        stage2_df <- data.frame(
          lab = as.numeric(cal_data[[lab_col]]),
          rf  = as.numeric(col_rf)
        ) %>% filter(!is.na(lab), !is.na(rf), is.finite(lab), is.finite(rf))

        if(nrow(stage2_df) < 3) {
          cat("    Too few colocation points for stage-2 fit of", param, "\n")
          next
        }

        stage2_model <- lm(lab ~ rf, data = stage2_df)
        stage2_r2    <- summary(stage2_model)$r.squared
        rf_r2        <- if(param %in% names(general_models)) general_models[[param]]$r_squared else NA_real_

        cat(sprintf("    %s  |  General RF R²: %.3f  |  Site-specific R²: %.3f\n",
                    param, rf_r2, stage2_r2))

        if(!is.na(rf_r2) && stage2_r2 <= rf_r2) {
          warning(sprintf(
            "⚠ Site-specific correction for %s [%s] does NOT improve on general RF (site-specific R²=%.3f ≤ general RF R²=%.3f). Recommend using rf_general instead.",
            site_name, param, stage2_r2, rf_r2
          ))
        }

        # Apply stage-2 to post-cutoff rows
        rf_vals_post <- corrected_data[[rf_col]][post_rows]
        corrected_data[[rf_col]][post_rows] <- coef(stage2_model)[1] +
                                               coef(stage2_model)[2] * rf_vals_post
        corrected_data$correction_method[post_rows] <- "rf_sitespecific"
      }
    } else if(any(post_rows) && is.null(cal_data)) {
      cat("  Warning: no colocation data found for", site_name, "— stage 2 skipped\n")
    }
  }

  cat("\nApplied calibration to", site_name,
      "(rf_general:", sum(!post_rows | !is_calibrated, na.rm=TRUE),
      "rows, rf_sitespecific:", sum(corrected_data$correction_method == "rf_sitespecific"), "rows)\n")

  corrected_data$date <- NULL
  return(corrected_data)
}

# ── Helper: apply the general RF model to a data frame ───────────────────
apply_rf_calibration <- function(sensor_data, general_models, site_name) {
  if(is.null(general_models) || length(general_models) == 0) {
    cat("  No general RF models available for", site_name, "\n")
    return(sensor_data)
  }
  corrected_data <- sensor_data
  target_params <- c("o3_ppb", "pm1_ugm3", "pm25_ugm3", "pm10_ugm3")

  for(param in target_params) {
    if(!param %in% names(general_models)) next

    pred_df <- build_rf_pred_df(sensor_data, param, type = "arduino")
    if(is.null(pred_df)) next

    preds <- predict_best_model(general_models[[param]], pred_df)
    corrected_data[[paste0("corrected_", param)]] <- preds
  }
  return(corrected_data)
}

# ── Helper: build a named predictor data frame for ranger ────────────────
# type = "arduino" uses arduino-naming (_arduino suffix, temp_c_arduino, humidity_perc_rh_arduino)
# type = "site"    uses site-naming    (_sensor suffix, temp_c_sensor,   humidity_perc_rh_sensor)
build_rf_pred_df <- function(data, param, type = "arduino") {
  if(type == "arduino") {
    sensor_col   <- paste0(param, "_arduino")
    temp_col     <- "temp_c_arduino"
    humidity_col <- "humidity_perc_rh_arduino"
  } else {
    sensor_col   <- paste0(param, "_sensor")
    temp_col     <- "temp_c_sensor"
    humidity_col <- "humidity_perc_rh_sensor"
  }

  if(!sensor_col %in% colnames(data)) return(NULL)

  # For the new models (MLR for Ozone, RF with Env for PM), 
  # RH and Temp are required predictors.
  if(temp_col %in% colnames(data) && humidity_col %in% colnames(data)) {
    df <- data.frame(
      x    = as.numeric(data[[sensor_col]]),
      temp = as.numeric(data[[temp_col]]),
      rh   = as.numeric(data[[humidity_col]])
    )
    colnames(df) <- c(sensor_col, temp_col, humidity_col)
    return(df)
  } else {
    # If missing, we can't apply the new models correctly.
    # Return a basic DF but warn or handle in calling function.
    # Given the requirement, we should ideally have them.
    cat(sprintf("  Warning: Missing environmental columns (%s/ %s) for %s prediction\n", 
                temp_col, humidity_col, param))
    df <- data.frame(x = as.numeric(data[[sensor_col]]))
    colnames(df) <- sensor_col
    return(df)
  }
}

# Example usage (uncomment and modify as needed):
# corrected_site_data <- apply_calibration_by_site(
#   site_sensor_data,
#   "SiteName",
#   general_models = general_models,
#   site_calibration_models = site_calibration_models,
#   colocation_data_list = NULL,   # or supply named list of colocation data frames
#   timestamp_col = "timestamp"
# )
```


# Print summary
```{r model-export}
if(!is.null(general_models) && length(general_models) > 0) {
  cat("\n\n")
  cat(paste0(rep("=", 80), collapse=""), "\n")
  cat("GENERAL RF CALIBRATION MODELS\n")
  cat(paste0(rep("=", 80), collapse=""), "\n")
  cat("Use these RF models for any site without site-specific calibration (Stage 1).\n\n")

  for(param in names(general_models)) {
    model_info <- general_models[[param]]
    cat("\n", toupper(param), ":\n")
    cat("  ", model_info$equation, "\n")
    cat("  Test R² =", round(model_info$r_squared, 4),
        ", Test RMSE =", round(model_info$rmse, 2),
        ", N =", model_info$n_points, "\n")
  }
  cat("\n")
  cat(paste0(rep("=", 80), collapse=""), "\n\n")
}

# Save unified model list for bmore_analysis_final.Rmd (General RF + Site-Specific Linear Refinement)
# This RDS file now contains both the Stage 1 (RF) and Stage 2 (Linear) coefficients.
if(!is.null(general_models) && length(general_models) > 0) {
  cat("\n", paste0(rep("=", 80), collapse=""), "\n")
  cat("PREPARING UNIFIED MODEL EXPORT FOR bmore_analysis_final.Rmd\n")
  cat(paste0(rep("=", 80), collapse=""), "\n")

  # 1. Prepare General RF Models (Stage 1)
  general_export <- list()
  for(param in names(general_models)) {
    mi <- general_models[[param]]
    param_name <- switch(param,
      "ozone_ppb" = "o3_ppb",   # Standardize to sensor naming
      "pm1_ugm3"  = "pm1_ugm3",
      "pm25_ugm3" = "pm25_ugm3",
      "pm10_ugm3" = "pm10_ugm3",
      param
    )

    general_export[[param_name]] <- list(
      model      = mi$model,        # renamed from ranger_model for simplicity
      model_type = mi$model_type,
      bias_value = mi$bias_value,
      r_squared  = mi$r_squared,   # test-set R²
      rmse       = mi$rmse,        # test-set RMSE
      importance = mi$importance,  # variable importance
      n_train    = mi$n_train,
      n_test     = mi$n_test,
      n_points   = mi$n_points
    )
    cat("  Included General RF model for", param_name, "\n")
  }

  # 2. Prepare Site-Specific Linear Refinements (Stage 2)
  # For each site in calibrated_sensors, we pre-calculate the linear shift:
  # lab_value ~ slope * general_rf_corrected + intercept
  cat("\n  Calculating site-specific Stage 2 refinements (Lab ~ RF_General)...\n")
  site_refinements <- list()
  
  if(exists("site_calibration_models") && length(site_calibration_models) > 0) {
    for(site in names(site_calibration_models)) {
      cal_data <- site_calibration_models[[site]]$combined_data
      if(is.null(cal_data)) next
      
      site_refinements[[site]] <- list()
      
      for(param in names(general_models)) {
        param_name <- switch(param, "o3_ppb"="ozone_ppb", "pm1_ugm3"="pm1_ugm3", 
                            "pm25_ugm3"="pm25_ugm3", "pm10_ugm3"="pm10_ugm3", param)
        lab_col <- paste0(param, "_vaisala")
        
        # Apply Stage 1 RF to the site's colocation data
        # Check alignment of columns in cal_data
        type_hint <- if(paste0(param, "_sensor") %in% names(cal_data)) "site" else "arduino"
        pred_df <- build_rf_pred_df(cal_data, param, type = type_hint)
        
        if(!is.null(pred_df) && lab_col %in% names(cal_data)) {
          # Apply General Model: MUST rename columns to match the general model's training names
          # (Param_arduino, humidity_perc_rh_arduino, temp_c_arduino)
          colnames(pred_df) <- c(paste0(param, "_arduino"), "humidity_perc_rh_arduino", "temp_c_arduino")
          
          # Use helper to apply Stage 1 (RF or LM)
          rf_general_preds <- predict_best_model(general_models[[param]], pred_df)
          
          stage2_df <- data.frame(lab = as.numeric(cal_data[[lab_col]]), rf = as.numeric(rf_general_preds)) %>%
            filter(!is.na(lab), !is.na(rf), is.finite(lab), is.finite(rf))
          
          if(nrow(stage2_df) >= 3) {
            stage2_model <- lm(lab ~ rf, data = stage2_df)
            
            # Extract coefficients
            s_int   <- as.numeric(coef(stage2_model)[1])
            s_slope <- as.numeric(coef(stage2_model)[2])
            
            # Handle case where Stage 1 was constant (bias-only), leading to NA slope 
            # We want to keep the intercept shift but have zero slope on the Stage 1 output
            if(is.na(s_slope)) s_slope <- 0
            if(is.na(s_int))   s_int   <- 0
            
            site_refinements[[site]][[param_name]] <- list(
              intercept = s_int,
              slope     = s_slope,
              r2        = summary(stage2_model)$r.squared,
              n_points  = nrow(stage2_df)
            )
            cat(sprintf("    %s [%s]: Stage 2 R² = %.3f (n=%d) | slope=%.3f, int=%.2f\n", 
                        site, param_name, summary(stage2_model)$r.squared, 
                        nrow(stage2_df), s_slope, s_int))
          }
        }
      }
    }
  }

  # 3. Save Unified RDS File
  coefficients_file <- "general_correction_coefficients.rds"
  export_list <- list(
    general = general_export,
    sites   = site_refinements,
    version = "2.0 (Unified RF+Linear)",
    export_date = Sys.Date()
  )
  saveRDS(export_list, file = coefficients_file)
  
  cat("\nUnified RF + Linear Refinement models saved to:", coefficients_file, "\n")
  cat("This file now supports the full two-stage correction in bmore_analysis_final.Rmd.\n")
  cat(paste0(rep("=", 80), collapse=""), "\n\n")
}
```

# RF Diagnostics — Predicted vs. Observed on Test Set
```{r rf-diagnostics, fig.width=6, fig.height=5}
# Diagnostic plots: RF-predicted vs. Vaisala-observed on the held-out test set.
# R² annotated on each plot so you can quickly judge model quality per pollutant.
if(exists("general_models") && length(general_models) > 0) {
  param_labels <- c(
    o3_ppm    = "Ozone (ppm)",
    pm1_ugm3  = "PM1 (μg/m³)",
    pm25_ugm3 = "PM2.5 (μg/m³)",
    pm10_ugm3 = "PM10 (μg/m³)"
  )

  for(param in names(general_models)) {
    mi <- general_models[[param]]
    arduino_col <- paste0(param, "_arduino")
    vaisala_col <- paste0(param, "_vaisala")

    # Reconstruct test set (must use SAME random seed and logic as training)
    set.seed(42)
    model_data <- all_data %>%
      filter(
        !is.na(.data[[arduino_col]]), !is.na(.data[[vaisala_col]]),
        is.finite(.data[[arduino_col]]), is.finite(.data[[vaisala_col]]),
        !is.na(temp_c_arduino), !is.na(humidity_perc_rh_arduino),
        is.finite(temp_c_arduino), is.finite(humidity_perc_rh_arduino)
      )
    
    n_train <- floor(0.8 * nrow(model_data))
    train_indices <- sample(seq_len(nrow(model_data)), n_train)
    test_df <- model_data[-train_indices, ]

    if(nrow(test_df) < 3) next

    pred_df <- build_rf_pred_df(test_df, param, type = "arduino")
    if(is.null(pred_df)) next

    # Predict using fallback-safe helper
    test_preds <- predict_best_model(mi, pred_df)
    r2_label   <- sprintf("R² = %.3f", mi$r_squared)

    diag_df <- data.frame(
      observed  = test_df[[vaisala_col]],
      predicted = test_preds
    )

    p <- ggplot(diag_df, aes(x = observed, y = predicted)) +
      geom_point(alpha = 0.5, colour = "steelblue", size = 1.5) +
      geom_abline(slope = 1, intercept = 0, colour = "firebrick", linetype = "dashed") +
      annotate("text", x = -Inf, y = Inf, label = r2_label,
               hjust = -0.1, vjust = 1.5, size = 4.5, colour = "firebrick") +
      labs(
        title    = paste0("Calibration diagnostics: ", param_labels[param]),
        subtitle = "Held-out test set (20%)",
        x        = paste("Observed Vaisala", param_labels[param]),
        y        = paste("Predicted", param_labels[param])
      ) +
      theme_minimal()

    print(p)

    # Variable importance bar chart (only for RF)
    if(!is.null(mi$importance)) {
      vi_df <- data.frame(
        variable   = names(mi$importance),
        importance = as.numeric(mi$importance)
      ) %>% arrange(desc(importance))

      p_vi <- ggplot(vi_df, aes(x = reorder(variable, importance), y = importance)) +
        geom_col(fill = "steelblue") +
        coord_flip() +
        labs(
          title = paste0("Variable importance: ", param),
          x = NULL, y = "Impurity-based importance"
        ) +
        theme_minimal()

      print(p_vi)
    }
  }
} else {
  cat("RF diagnostics: no general_models available yet.\n")
}
```

# Apply Corrections
```{r apply-corrections}
# Applies the general RF calibration to sensor data.
# Input: data frame with Arduino measurements + humidity + temp (all_data columns)
# Output: same data frame with new "corrected_*" columns AND correction_method column.
apply_general_calibration <- function(all_data, general_models) {
  if(is.null(general_models) || length(general_models) == 0) {
    cat("No general calibration models available for correction.\n")
    return(all_data)
  }

  corrected_data <- all_data

  for(param in names(general_models)) {
    model_info  <- general_models[[param]]
    corrected_col <- paste0("corrected_", param)

    pred_data <- build_rf_pred_df(all_data, param, type = "arduino")
    if(is.null(pred_data)) {
      cat("Skipping", param, "— sensor column not found\n")
      next
    }

    # Predict using fallback-safe helper
    corrected_data[[corrected_col]] <- predict_best_model(model_info, pred_data)

    cat(sprintf("Applied %s calibration for %s (test R² = %.3f, test RMSE = %.2f)\n",
                toupper(model_info$model_type), param, model_info$r_squared, model_info$rmse))
  }

  # Tag all rows with the general method identifier
  corrected_data$correction_method <- "general_calibration"

  return(corrected_data)
}

# Apply corrections
if(exists("general_models") && length(general_models) > 0) {
  corrected_data <- apply_general_calibration(all_data, general_models)

  cat("\nCorrections applied successfully!\n")
  cat("Corrected columns:", paste(grep("corrected_", colnames(corrected_data), value = TRUE), collapse = ", "), "\n")
  cat("correction_method values:", paste(unique(corrected_data$correction_method), collapse = ", "), "\n")
} else {
  cat("No general calibration models available for correction.\n")
  corrected_data <- all_data
}
```

#Time Series Plots

```{r time-series-visualization, echo=TRUE, message=FALSE, warning=FALSE}
# Simple time series comparison
create_simple_plots <- function() {
  
  if(!exists("corrected_data")) {
    cat("Corrected data not available.\n")
    return(NULL)
  }
  
  # Parameters to visualize
  parameters <- list(
    "o3_ppb" = "Ozone (ppb)",
    "pm1_ugm3" = "PM1 (μg/m³)", 
    "pm25_ugm3" = "PM2.5 (μg/m³)",
    "pm10_ugm3" = "PM10 (μg/m³)"
  )
  
  for(param in names(parameters)) {
    param_label <- parameters[[param]]
    vaisala_col <- paste0(param, "_vaisala")
    corrected_col <- paste0("corrected_", param)
    
    arduino_col <- paste0(param, "_arduino")
    
    # Ensure all required columns exist
    if(!(vaisala_col %in% colnames(corrected_data) &&
         arduino_col %in% colnames(corrected_data) &&
         corrected_col %in% colnames(corrected_data))) next
    
    # Build plotting data with raw timestamps in UTC
    plot_data <- corrected_data %>%
      dplyr::select(timestamp, all_of(c(vaisala_col, arduino_col, corrected_col))) %>%
      mutate(timestamp = parse_to_utc(timestamp)) %>%
      filter(if_any(all_of(c(vaisala_col, arduino_col, corrected_col)), ~ !is.na(.x)))
    
    if(nrow(plot_data) < 3) next
    
    # Simple time-series plot with three series and proper legend
    p <- ggplot(plot_data, aes(x = timestamp)) +
      geom_line(aes(y = .data[[arduino_col]], color = "Arduino original", linetype = "Arduino original"),
                alpha = 0.7, linewidth = 0.6) +
      geom_line(aes(y = .data[[corrected_col]], color = "Arduino corrected", linetype = "Arduino corrected"),
                alpha = 0.7, linewidth = 0.6) +
      geom_line(aes(y = .data[[vaisala_col]], color = "Vaisala reference", linetype = "Vaisala reference"),
                alpha = 0.7, linewidth = 0.6) +
      labs(
        title = paste("Time Series:", param_label),
        x = "Timestamp",
        y = param_label,
        color = NULL,
        linetype = NULL
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_color_manual(values = c(
        "Arduino original" = "red",
        "Arduino corrected" = "blue",
        "Vaisala reference" = "green"
      )) +
      scale_linetype_manual(values = c(
        "Arduino original" = "dashed",
        "Arduino corrected" = "dotted",
        "Vaisala reference" = "solid"
      ))
    
    print(p)
  }
}

create_simple_plots()
```

#Comparisons
```{r paired-comparisons, echo=TRUE, message=FALSE, warning=FALSE}
# Paired comparisons: (1) Arduino original vs corrected, (2) Corrected vs Vaisala
create_comparison_plots <- function() {
  
  if(!exists("corrected_data")) {
    cat("Corrected data not available.\n")
    return(NULL)
  }
  
  parameters <- list(
    "o3_ppb" = "Ozone (ppb)",
    "pm1_ugm3" = "PM1 (μg/m³)",
    "pm25_ugm3" = "PM2.5 (μg/m³)",
    "pm10_ugm3" = "PM10 (μg/m³)"
  )
  
  for(param in names(parameters)) {
    param_label <- parameters[[param]]
    arduino_col <- paste0(param, "_arduino")
    corrected_col <- paste0("corrected_", param)
    vaisala_col <- paste0(param, "_vaisala")
    
    needed_cols <- c("timestamp", arduino_col, corrected_col, vaisala_col)
    if(!all(needed_cols %in% colnames(corrected_data))) next
    
    plot_data <- corrected_data %>%
      dplyr::select(all_of(needed_cols)) %>%
      mutate(timestamp = parse_to_utc(timestamp)) %>%
      filter(if_any(all_of(needed_cols[-1]), ~ !is.na(.x)))
    
    if(nrow(plot_data) < 3) next
    
    # (1) Arduino original vs Arduino corrected
    p1 <- ggplot(plot_data, aes(x = timestamp)) +
      geom_line(aes(y = .data[[arduino_col]], color = "Arduino original", linetype = "Arduino original"),
                alpha = 0.8, linewidth = 0.6) +
      geom_line(aes(y = .data[[corrected_col]], color = "Arduino corrected", linetype = "Arduino corrected"),
                alpha = 0.8, linewidth = 0.6) +
      labs(
        title = paste(param_label, "— Original vs Corrected"),
        x = "Timestamp",
        y = param_label,
        color = NULL,
        linetype = NULL
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_color_manual(values = c(
        "Arduino original" = "red",
        "Arduino corrected" = "blue"
      )) +
      scale_linetype_manual(values = c(
        "Arduino original" = "dashed",
        "Arduino corrected" = "solid"
      ))
    print(p1)
    
    # (2) Arduino corrected vs Vaisala reference
    p2 <- ggplot(plot_data, aes(x = timestamp)) +
      geom_line(aes(y = .data[[corrected_col]], color = "Arduino corrected", linetype = "Arduino corrected"),
                alpha = 0.8, linewidth = 0.6) +
      geom_line(aes(y = .data[[vaisala_col]], color = "Vaisala reference", linetype = "Vaisala reference"),
                alpha = 0.8, linewidth = 0.6) +
      labs(
        title = paste(param_label, "— Corrected vs Vaisala"),
        x = "Timestamp",
        y = param_label,
        color = NULL,
        linetype = NULL
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_color_manual(values = c(
        "Arduino corrected" = "blue",
        "Vaisala reference" = "green"
      )) +
      scale_linetype_manual(values = c(
        "Arduino corrected" = "dotted",
        "Vaisala reference" = "solid"
      ))
    print(p2)
  }
}

create_comparison_plots()
```


# Barkjohn (EPA) PM2.5 Comparison
```{r barkjohn-comparison, fig.width=10, fig.height=6}
# -------------------------------------------------------------------------
# BARKJOHN (EPA) PM2.5 BENCHMARKING
# -------------------------------------------------------------------------
# This section implements the standard US EPA correction for PM2.5 
# (Barkjohn et al., 2021) to benchmark against our custom Vaisala-based RF.
# Formula: PM2.5_Barkjohn = 0.524 * PM2.5_raw - 0.0862 * RH + 5.75
# -------------------------------------------------------------------------

if(exists("all_data_hourly") && nrow(all_data_hourly) > 0) {
  cat("\n=== Benchmarking against Barkjohn (EPA) PM2.5 Correction ===\n")
  
  # 1. Apply Barkjohn correction
  # Using CF=1 (which is the default assumption for the 0.524 factor)
  barkjohn_data <- all_data_hourly %>%
    mutate(
      pm25_barkjohn = 0.524 * pm25_ugm3_arduino - 0.0862 * humidity_perc_rh_arduino + 5.75
    ) %>%
    filter(!is.na(pm25_barkjohn), !is.na(pm25_ugm3_vaisala))
  
  # 2. Get our RF correction for comparison (if available)
  # We can use the Stage 1 general model for this benchmark
  if(!is.null(general_models) && "pm25_ugm3" %in% names(general_models)) {
    pred_df <- build_rf_pred_df(barkjohn_data, "pm25_ugm3", type = "arduino")
    if(!is.null(pred_df)) {
      barkjohn_data$pm25_rf_custom <- predict_best_model(general_models[["pm25_ugm3"]], pred_df)
    }
  }
  
  # 3. Calculate Performance Metrics
  metrics <- barkjohn_data %>%
    summarise(
      barkjohn_r2 = 1 - sum((pm25_ugm3_vaisala - pm25_barkjohn)^2) / sum((pm25_ugm3_vaisala - mean(pm25_ugm3_vaisala))^2),
      barkjohn_rmse = sqrt(mean((pm25_ugm3_vaisala - pm25_barkjohn)^2)),
      rf_custom_r2 = if("pm25_rf_custom" %in% colnames(.)) 
                       1 - sum((pm25_ugm3_vaisala - pm25_rf_custom)^2) / sum((pm25_ugm3_vaisala - mean(pm25_ugm3_vaisala))^2) 
                     else NA_real_,
      rf_custom_rmse = if("pm25_rf_custom" %in% colnames(.)) 
                         sqrt(mean((pm25_ugm3_vaisala - pm25_rf_custom)^2)) 
                       else NA_real_
    )
  
  cat("PM2.5 Performance Table:\n")
  print(round(metrics, 4))
  
  # 4. Visualization
  plot_long <- barkjohn_data %>%
    dplyr::select(timestamp, pm25_ugm3_vaisala, pm25_barkjohn, 
           any_of("pm25_rf_custom")) %>%
    pivot_longer(cols = -timestamp, names_to = "Method", values_to = "PM25") %>%
    mutate(Method = factor(Method, 
                          levels = c("pm25_ugm3_vaisala", "pm25_rf_custom", "pm25_barkjohn"),
                          labels = c("Vaisala (Reference)", "Custom RF (Stage 1)", "Barkjohn (EPA)")))
  
  p_comp <- ggplot(plot_long, aes(x = timestamp, y = PM25, color = Method)) +
    geom_line(alpha = 0.7) +
    labs(
      title = "PM2.5 Correction Benchmark: Custom RF vs Barkjohn (EPA)",
      subtitle = sprintf("Vaisala-based RF R²: %.3f | Barkjohn R²: %.3f", 
                         metrics$rf_custom_r2, metrics$barkjohn_r2),
      x = "Timestamp",
      y = "PM2.5 (μg/m³)",
      color = "Method"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_color_manual(values = c("Vaisala (Reference)" = "black", 
                                "Custom RF (Stage 1)" = "blue", 
                                "Barkjohn (EPA)" = "red"))
  
  print(p_comp)
  
} else {
  cat("Insufficient data in all_data_hourly for Barkjohn comparison.\n")
}
```

